### 2014-10-28 intro1.tex

Many real-world web services --- such as those built and maintained by
Amazon, Facebook, Google, Twitter, etc. --- replicate application state and
logic across multiple \emph{replicas} within and across data
centers. Replication is intended not only to improve application throughput
and reduce user-perceived latency, but also to tolerate partial failures
without compromising overall service availability. Traditionally programmers
have relied on \emph{strong consistency} guarantees such as
linerarizability~\cite{Herlihy1990} or
serializability~\cite{Serializability} in order to build correct
applications.  While strong consistency is an easily stated property, it
masks the reality underlying large-scale distributed systems with respect to
non-uniform latency (i.e., availability) and network
partitions~\cite{Brewer2000,Gilbert}. Indeed, modern web services, which aim
to provide an "always on" experience, overwhelmingly favour availability and
partition tolerance over strong consistency. To this end, several \emph{weak
  consistency} models such as eventual consistency, causal consistency,
session guarantees, and timeline consistency have been proposed.

Under weak consistency, the developer needs to be aware of concurrent
conflicting updates, and has to pay careful attention to avoid unwanted
inconsistencies (e.g., negative balances in a bank account, or having an
item appear in a shopping cart after it has been
removed~\cite{Dynamo}). Oftentimes, the inconsistency leaks from the
application and is witnessed by the user.  Ultimately, the developer must
decide the consistency level appropriate for a particular operation; this is
understandably an error-prone process requiring intricate knowledge of both
the application as well as the semantics and implementation of the
underlying data store, which typically have only informal
descriptions~\cite{}.  Nonetheless, picking the correct consistency level is
critical not only for correctness but also for scalability of the
application. While choosing a weaker consistency level than required may
introduce program errors and anomalies, choosing a stronger one than
necessary can negatively impact program scalability.

@@

Weak consistency also hinders compositional reasoning about programs.  While
an application might be naturally expressed in terms of well-understood and
expressive data types such as maps, trees, queues, or graphs,
geo-distributed stores typically only provide a minimal set of data types
with in-built conflict resolution strategies such as last-writer-wins (LWW)
registers, counters, and sets~\cite{Cassandra,DynamoDB}.  Furthermore, while
traditional database systems enable composability through transactions,
geo-distributed stores typically lack unrestricted transactional access to
the data.  Working in this environment thus requires application state to be
suitably coerced to function using only the capabilities of the store.

To address these issues, we describe \name, a declarative programming model
and implementation for eventually consistent geo-distributed data
stores. The key novelty of \name is an expressive \emph{contract language}
to declare and verify fine-grained application-level consistency
properties. The programmer uses the contract language to axiomatically
specify the set of legal executions allowed over the replicated data
type. Contracts are constructed using primitive consistency relations such
as \emph{visibility} and \emph{session order} along with standard logical
and relational operators. A \emph{contract enforcement system} automatically
maps operations over the datatype to a particular consistency level
available on the store, and provably validates the correctness of the
mapping.  The paper makes the following contributions:

\begin{itemize}
\setlength{\itemsep}{2pt}
\item We introduce \name, a language for declaratively describing eventually
	consistent programs that manipulate replicated data types. Contracts are used
	to specify fine-grained application-level consistency properties, and are
	analyzed to assign the most efficient and sound store consistency level to
	the corresponding operation.
\item \name supports coordination-free transactions over arbitrary datatypes.
	We extend our contract language to express fine-grained transaction isolation
	guarantees, and utilize the contract enforcement system to automatically
	assign the correct and most efficient isolation level for a transaction.
\item We provide metatheory that certifies the soundness of our contract
	enforcement system, and ensures that an operation is only executed if the
	required conditions on consistency are met.
\item An implementation of \name as a transparent shim layer over
	Cassandra~\cite{Cassandra}, a well-known general-purpose data store.  An
	extensive experimental evaluation over a collection of real-world
	applications, including a Twitter-like micro-blogging site and an eBay-like
	auction site illustrates the practicality of our approach.
\end{itemize}

>>>>

Many real-world web services --- such as those built and maintained by
Amazon, Facebook, Google, Twitter, etc. --- replicate application state and
logic across multiple \emph{replicas} within and across data
centers. Replication is intended not only to improve application throughput
and reduce user-perceived latency, but also to tolerate partial failures
without compromising overall service availability. Traditionally programmers
have relied on \emph{strong consistency} guarantees such as
linerarizability~\cite{Herlihy1990} or
serializability~\cite{Serializability} in order to build correct
applications. While strong consistency is an easily stated property, it
masks the reality underlying large-scale distributed systems with respect to
non-uniform latency (i.e., availability) and network
partitions~\cite{Brewer2000,Gilbert}.  Indeed, modern web services, which
aim to provide an "always on" experience, overwhelmingly favour availability
and partition tolerance over strong consistency. Several \emph{weak
  consistency} models such as eventual consistency, causal consistency,
session guarantees, and timeline consistency have consequently been
proposed.

Under weak consistency, the developer needs to be aware of concurrent
conflicting updates, and has to pay careful attention to avoid unwanted
inconsistencies (e.g., negative balances in a bank account, or having an
item appear in a shopping cart after it has been
removed~\cite{Dynamo}). Oftentimes, the inconsistency leaks from the
application and is witnessed by the user.  Ultimately, the developer must
decide the consistency level appropriate for a particular operation; this is
understandably an error-prone process requiring intricate knowledge of both
the application as well as the semantics and implementation of the
underlying data store.  More often than not, only informal descriptions are
available~\cite{}.  Picking the correct consistency level is critical not
only for correctness but also for scalability of the application. While
choosing a weaker consistency level than required may introduce program
errors and anomalies, choosing a stronger one than necessary level can
negatively impact program scalability.

@@

Moreover, weak consistency hinders compositional reasoning about programs.
While an application might be naturally expressed in terms of
well-understood and expressive data types such as maps, trees, queues, or
graphs, geo-distributed stores typically only provide a minimal set of data
types with in-built conflict resolution strategies such as last-writer-wins
(LWW) registers, counters, and sets~\cite{Cassandra,DynamoDB}.  Furthermor,
while traditional database systems enable composability through
transactions, existing geo-distributed stores lack unrestricted
transactional access to the data.  Working with geo-distributed stores thus
requires application state to be suitably coerced to work within the
capabilities of the store.

To address these issues, we describe \name, a declarative programming model
and implementation for eventually consistent stores. The key novelty of
\name\ is an expressive \emph{contract language} to declare and verify
fine-grained application-level consistency properties.  The programmer uses
the contract language to axiomatically describe the set of legal executions
allowed over the replicated data type.  Contracts are constructed using
primitive consistency relations such as \emph{visibility} and \emph{session
  order} along with standard logical and relational operators.  A
\emph{contract enforcement system} automatically maps each operation to a
particular consistency level available on the store.  In addition, the
system statically reports an error if the given application-level
consistency property cannot be satisfied by the store.

The paper makes the following contributions:

\begin{itemize}

\item We introduce \name, a language for declaratively describing eventually
  consistent programs that manipulate replicated data types. Contracts are
  used to specify fine-grained application-level consistency properties, and
  are analyzed to assign the most efficient and sound store consistency
  level to the corresponding operation.

\item \name supports coordination-free transactions over arbitrary
  datatypes.  We extend our contract language to express fine-grained
  transaction isolation guarantees, and utilize the contract enforcement
  system to automatically assign the correct and most efficient isolation
  level for a transaction.

\item Metatheory that certifies the soundness of our contract enforcement
  system, and ensures that an operation is only executed if the required
  conditions on consistency are met is also given.

\item \name is implemented as a shim layer over (and importantly without
  modifying) Cassandra, a well-known general-purpose data store.  An
  extensive experimental evaluation over a collection of real-world
  applications, including a Twitter-like micro-blogging site and an
  eBay-like auction site illustrates the practicality of our approach.

\end{itemize}
